Wrinkling of a confined porous layer

### 2014-09-22 plis_beamer.tex
Remove old speculations and partial scalings. Better present Poiseuille. Use permeability instead of pore size. What can we change experimentally.

\bigskip
\begin{itemize}
\item solvent must flow through gel
\begin{description}[$\ell_p$]
\item[$v$] gel velocity = -volumic flux $\simeq \SI{0.2}{\micro\metre\per\second}$
\item[$\ell_p$] pore size $ \approx \SI{4}{\micro\metre}$
\end{description}
\item $\sigma$ is a Darcy pressure
\[\sigma = \Delta P \sim \eta h v/\ell_p^2\]
\item poroelastic length
\[ \lambda^* \sim \left(\frac{B \ell_p^2}{\eta v h}\right)^\frac{1}{3} \simeq \SI{2}{\milli\metre}\]
\end{itemize}
}{
\tikzsetnextfilename{alts_bottom_cloque}
\begin{tikzpicture}
\begin{axis}[
	width=\textwidth,
	height=0.6\textwidth,
	xmin=0, xmax=1000, xtick={0,250, 500, 750}, xlabel={position (\si{\micro\metre})},
	ymin=0, ymax=50, ylabel={altitude (\si{\micro\metre})},
	cycle list name=earthy,
	no marks,
	]
	\draw[ultra thick, Accent2] (axis cs:580,0) -- (axis cs:580,50);
	\foreach \x in {2,3,..., 8}
		\addplot table[y index=\x]{alts_bottom_cloque.txt};
\end{axis}
\end{tikzpicture}
\begin{itemize}
\item $\lambda$ frozen at ceiling touch 
\item $A=H\equiv e-h$
\item $\lambda \sim \epsilon^{-3/7} H \sim e$
	%\[ \epsilon \sim \left(\frac{\lambda}{e}\frac{e}{H}\right)^{-7/3} \simeq 0.6\% \]
\end{itemize}
}

@@

\begin{columns}
\column{0.83\textwidth}
\begin{itemize}
\item The same $\Delta P$ can be relaxed by a Poiseuille flow
\[\Delta P \sim 12\eta\frac{\lambda}{H^2} u \sim 12\eta\frac{\lambda^2}{H^3} v\]
\item For the same wavelength
\hfill$\displaystyle \frac{v_\text{Darcy}}{v_\text{Poiseuille}} \sim \frac{\ell_p^2\lambda^2}{hH^3} \approx 30$
\item Flow through the pores dominates
\end{itemize}
\column{0.17\textwidth}
\includegraphics[width=\textwidth]{Jean-Leonard-Marie_Poiseuille.jpg}\\
\hfill\structure{1797-1869}
\end{columns}
\end{frame}

@@

%\tikzset{external/force remake}
\begin{frame}{Cell thickness}
\tikzsetnextfilename{thickness}
\begin{tikzpicture}[mark size={0.3em}]
	\matrix[matrix of nodes, inner sep=0, column sep=0.01\columnwidth, row sep=0.01\columnwidth, ampersand replacement=\&] (m) {
		\includegraphics[width=0.245\columnwidth]{pattern_50um.jpg}\&
		\includegraphics[width=0.245\columnwidth]{pattern_100um.jpg}\&
		\includegraphics[width=0.245\columnwidth]{pattern_250um.jpg}\&
		\includegraphics[width=0.245\columnwidth]{pattern_450um.jpg}\\
	};
	\node[inner sep=0, below left=0.01\columnwidth and 0 of m.south east] (x800) {\includegraphics[width=0.49\columnwidth]{pattern_800um_x1.jpg}};
	\draw[ultra thick, Accent2, <->] (x800.south west) ++(0.196\columnwidth, 0.05\columnwidth) -- ++(0.145\columnwidth,0) node[midway,above]{$\lambda$};
	\draw[line width=0.2em,Main](x800.north east) -- ++(-0.217\columnwidth,0) node[midway,below] {\SI{5}{\milli\metre}};

	\begin{axis}[%
		name=g,
		anchor=left of north west,
		at={($(m.south west)+(0,-0.01\columnwidth)$)},
		scale only axis,
		width=0.5\textwidth-3em,
		height=0.25\columnwidth-2em,
		xmin=0,xmax=1,xlabel={$e$ (\si{\milli\metre})},
		xtick={0,0.4,...,0.9},
		ymin=0,ylabel={$\lambda$ (\si{\milli\metre})},
		]
		\addplot[no marks, domain=0:3] {4.23*x};
		\begin{scope}[inner sep=0, minimum size=0.8em]
			\node[circle, fill=Accent1] at (axis cs:0.05,0.261){};
			\node[circle, fill=white, draw=Accent1] at (axis cs:0.1,0.458) {};
			\node[regular polygon, regular polygon sides=3, fill=Accent1] at (axis cs:0.250,0.916){};
			\node[regular polygon, regular polygon sides=3, fill=white, draw=Accent1] at (axis cs:0.450,2.075){};
			\node[fill=Accent1] at (axis cs:0.800,3.333){};
		\end{scope}
	\end{axis}
	\begin{scope}[inner sep=0, minimum size=0.8em, transform canvas={xshift=0.5em, yshift=-0.5em}]
		\node[circle, fill=Accent1] at (m-1-1.north west) {};
		\node[circle, fill=white, draw=Accent1] at (m-1-2.north west){};
		\node[regular polygon, regular polygon sides=3, fill=Accent1] at (m-1-3.north west){};
		\node[regular polygon, regular polygon sides=3, fill=white, draw=Accent1] at (m-1-4.north west) {};
		\node[fill=Accent1] at (x800.north west) {};
	\end{scope}
\end{tikzpicture}
\end{frame}
%\tikzset{external/force remake=false}

%\tikzset{external/force remake}
\begin{frame}{Surprising viscosity influence}
\begin{columns}
\column{0.5\textwidth}
\[ \lambda^* \sim \left(\frac{B \ell_p^2}{\eta v h}\right)^\frac{1}{3}\]
\tikzsetnextfilename{viscosity_dependence}
\begin{tikzpicture}
	\begin{loglogaxis}[%
		width=\textwidth,
		height=0.8\textwidth,
		xmin=0.6,xmax=11,xlabel={viscosity (\si{\milli\pascal\second})},
		xtick={0.7, 1, 2, 5, 10}, xticklabels={$0.7$, $1$, $2$, $5$, $10$},
		ytick={1,2,3,4}, yticklabels={$1$,$2$,$3$,$4$},
		ymin=1, ymax=5, ylabel={$\lambda$ (\si{\milli\metre})},
		mark size={0.3em},
		legend style={legend pos=south east},
		]
		\addplot+[no marks, domain=0.7:10, black, forget plot] {1.7*x^0.4} node[midway, below] {$\eta^{0.4}$};
		\addplot+[only marks,Accent1, mark=o] table{eta_lambda_eau_pure.txt};
		\addplot+[only marks,Accent2, mark=square] table{eta_lambda_eau_glycerol.txt};
		\legend{water, water-glycerol};
	\end{loglogaxis}
\end{tikzpicture}

\column{0.5\textwidth}
\begin{itemize}
\item $\lambda$ increases with $\eta$
\item Can be explained only by a hidden dependence of the viscosity on $B$, $\ell_p$, \ldots 
\item[$\Rightarrow$] Viscosity influences gel formation.
\end{itemize}
%If $\ell_p \sim \eta^{-\alpha}$ and $B\sim \mathrm{k_B}T \left(\frac{h}{\ell}\right)^3$, then
%\[ \lambda^* \sim \eta^\frac{\alpha-1}{3} \]
%\[\alpha\simeq 2.2\]
\end{columns}
\end{frame}

@@

\structure{To do:} Measure everything to plot $\lambda_\text{theory}$ vs $\lambda_\text{experiment}$
\end{frame}

>>>>

\bigskip
\begin{itemize}
\item solvent must flow through gel
\begin{description}[$\ell_p$]
\item[$v$] gel velocity = -volumic flux $\simeq \SI{0.2}{\micro\metre\per\second}$
\item[$\alpha$] permeability $ \approx \SI{6e-14}{\square\metre}$
\end{description}
\item $\sigma$ is a Darcy pressure
\[\sigma = \Delta P \sim \eta h v/\alpha\]
\item poroelastic length
\[ \lambda^* \sim \left(\frac{B \alpha}{\eta v h}\right)^\frac{1}{3} \simeq \SI{0.5}{\milli\metre}\]
\end{itemize}
}{
\tikzsetnextfilename{alts_bottom_cloque}
\begin{tikzpicture}
\begin{axis}[
	width=\textwidth,
	height=0.6\textwidth,
	xmin=0, xmax=1000, xtick={0,250, 500, 750}, xlabel={position (\si{\micro\metre})},
	ymin=0, ymax=50, ylabel={altitude (\si{\micro\metre})},
	cycle list name=earthy,
	no marks,
	]
	\draw[ultra thick, Accent2] (axis cs:580,0) -- (axis cs:580,50);
	\foreach \x in {2,3,..., 8}
		\addplot table[y index=\x]{alts_bottom_cloque.txt};
\end{axis}
\end{tikzpicture}
\begin{itemize}
\item $\lambda$ frozen at ceiling touch 
\item Amplitude $A=e-h$
\item $\epsilon = \left(\frac{e-h}{\lambda^*}\right)^\frac{7}{4}$
\item $\lambda \sim \left(e-h\right)^\frac{1}{4} \left(\lambda^*\right)^\frac{3}{4}$
	%\[ \epsilon \sim \left(\frac{\lambda}{e}\frac{e}{H}\right)^{-7/3} \simeq 0.6\% \]
\end{itemize}
}

@@

\begin{columns}
\column{0.17\textwidth}
\includegraphics[width=\textwidth]{Jean-Leonard-Marie_Poiseuille.jpg}\\
\hfill\structure{1797-1869}

\column{0.80\textwidth}
The same $\Delta P$ can be relaxed by a Poiseuille flow
\begin{align*}
\Delta P &\sim 12\eta\frac{\lambda}{H^2} u \sim 12\eta\frac{\lambda^2}{H^3} v\\
\lambda^* &= \left(\frac{B H^2}{12\eta\lambda}\right)^{1/3}
\end{align*}
\textit{\footnotesize Huang \& Suo, J. Appl. Phys. (2002)}
%\item For the same wavelength
%\hfill$\displaystyle \frac{v_\text{Darcy}}{v_\text{Poiseuille}} \sim \frac{\ell_p^2\lambda^2}{hH^3} \approx 30$
%\item Flow through the pores dominates

\end{columns}
\end{frame}

@@

\begin{frame}{Poiseuille vs Darcy}
\begin{columns}
\column{0.5\textwidth}
\begin{block}{Observable wavelengths}
Same magnitude (typical sample)
\begin{align*}
\lambda_D &\sim \Upsilon^\frac{1}{4} h^\frac{1}{2} \alpha^\frac{1}{4}\simeq \SI{0.23}{\milli\metre}\\
\lambda_P &\sim \Upsilon^\frac{1}{6} h^\frac{1}{2} H^\frac{1}{2}\simeq \SI{0.25}{\milli\metre}
\end{align*}
\end{block}
with $\Upsilon = \frac{E(e-h)}{\eta v}\simeq 10^{6-8}$
\column{0.5\textwidth}
Limiting case:
\begin{itemize}
\item Same growth speed
\item Same wavelength
\item[$\Rightarrow$] $H^* = \Upsilon^\frac{1}{6} \alpha^\frac{1}{2}\simeq \SI{6}{\micro\metre}$
\end{itemize}
Darcy can be faster than Poiseuille when $H<H^*$.
\end{columns}
\begin{block}{What can we change experimentally?}
$H$ and $v$ are measured a posteriori but cannot be controlled.
\begin{description}[leftmargin=8em]
\item[Geometry] $e$, $h$
\item[Composition] $E$, $\alpha$
\item[Solvent] $\eta$ \alert{but also $E$ and $\alpha$}
\end{description}
\end{block}
\end{frame}

@@

May be explained by ``viscoelastic phase separation'' arguments
\end{frame}

\begin{frame}{Darcy never beats Poiseuille}
